A Comparative study of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-15 latex/paper.tex
Made correction to comment + Modified Bids App to talk about singularity

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
provides in-memory computing, data locality, and lazy evaluation. Spark's main data
structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
and a parallel collection of data elements. Dataframes, Spark's API for tabular data,
is an overlay of RDDs with similar performance features \VHS{It's probably fine, but
the SparkSQL Engine available for Scala and Java do bring an added peformance
value}.\VHS{Spark has another API (Datasets). i know you used 'popular' for
DataFrames, but it's not like Spark has a plethora of APIs. I think the APIs should
be mentioned after the API languages such that you only list the APIs relevant to
Python (RDD and Dataframes)} Spark's core is written in Scala, but it also has a
Java, an R, and a Python API called PySpark. \MD{Sentence is ambiguous? could think
that all three APIs are named PySpark} We used PySpark, given the popularity of
Python in neuroscience, although the serialization cost of Python objects to Scala
might be a substantial source of overhead. We used Spark's default scheduler, Spark
standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
a process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
in turn requests workers to the master and dispatches tasks to them. A job is divided
into stages to be excuted in a different process onto the workers. Each stage's
operation is represented as a high-level task in the computation graph. The Spark
standalone scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
Processing jobs in order helps in increasing the page hit rate. As our applications
only have one job, the LIFO (Last-In-First-Out) policy used for the stages is more
pertinent \MD{Find better phrasing} \TG{refer to the paper (\MD{Not sure which one})}
\VHS{I wouldn't explicitly state LIFO here. It's a bit complicated to be frank.
Basically both Spark and Dask opt to complete a chain of dependencies rather than
start a bunch of ``new chains''. And in this context, LIFO is great for the page hit
rate, so you might want to reword the prior sentence as well}. The Spark standalone
scheduler has two execution modes: client mode, where the driver runs in a dedicated
process on the cluster, and cluster mode, where the driver runs in a worker. We used
client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.

@@

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It use multi-process to enable
parallelism. It offers a programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. An
\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the processing
of large arrays. It provides a distributed clone of the popular NumPy library. A
\href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.
\VHS{maybe mention which one use threads and which use processes}

@@

While Incrementation and Histogram were implemented directly in Python,
BIDS app example requires an external library and is distributed as a
Docker container image (bids/base\_fsl on DockerHub). To avoid remote file
transfers from DockerHub, we preloaded our OpenStack instances with the
Docker image, and started the Docker engine at startup.
\subsection{Experiments}

@@

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead across all application tasks. 
We measured the overhead by \TG{\ldots, maybe add a figure}.

>>>>

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine.
Spark's core is written in Scala, but it also has a Java, an R, and a Python API
called PySpark. We used PySpark, given the popularity of Python in neuroscience,
although the serialization cost of Python objects to Scala might be a substantial
source of overhead. Spark provides in-memory computing, data locality, and lazy
evaluation. PySpark's fundamental data structure, the Resilient Distributed Dataset
(RDD)~\cite{RDD}, is a fault-tolerant and a parallel collection of data elements.
Dataframes, PySpark's API for tabular data, is an overlay of RDDs with similar
performance features. We used Spark's default scheduler, Spark standalone, although
YARN~\cite{vavilapalli2013apache} and Mesos~\cite{hindman2011mesos} are also
available. In the Spark standalone scheduler, a process called the \emph{master}
coordinates resources provisioned by \emph{worker} processes on the cluster. The
application is submitted to the \emph{driver} that in turn requests workers to the
master and dispatches tasks to them. A job is divided into stages to be excuted in a
different process onto the workers. Each stage's operation is represented as a
high-level task in the computation graph. The Spark standalone scheduler uses a FIFO
(First-In-First-Out) job scheduling policy. Processing jobs in order helps in
increasing the page hit rate. As our applications only have one job, the LIFO
(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't
explicitly state LIFO here. It's a bit complicated to be frank. Basically both Spark
and Dask opt to complete a chain of dependencies rather than start a bunch of ``new
chains''. And in this context, LIFO is great for the page hit rate, so you might want
to reword the prior sentence as well}. The Spark standalone scheduler has two
execution modes: client mode, where the driver runs in a dedicated process on the
cluster, and cluster mode, where the driver runs in a worker. We used client mode as
cluster mode is not available in PySpark. We used Apache Spark v2.4.0.

@@

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It use multi-process to enable
parallelism. It offers a programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. An
\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the processing
of large arrays. It provides a distributed clone of the popular NumPy library. A
\href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.

@@

While Incrementation and Histogram were implemented directly in Python, BIDS app
example requires an external library and is distributed as a Docker container image
(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
given its popularity for HPC and reproducibility, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
The image was preloaded on the nfs and singularity version 3.2.1\-1.el7 was installed
on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
export the path for fsl.

\subsection{Experiments}

@@

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead across all application tasks. 
We measured the overhead by \TG{\ldots, maybe add a figure}. \MD{Done but might need
adjustment depending on how the overhead is calculate (idle vs wasted)}
